---
title: "Project"
author: "Tianyi Xia, Zetian Zhao, Litao Zheng"
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%OS')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Species** =\> `Great Blue Heron`.

## Spatial Analysis

```{r}
# Load libraries
library(spatstat)
library(sf)
library(raster)
```

## Load the Covariates

```{r}
load("BC_Covariates.Rda")
ls()
```

## Summary

```{r}
summary(DATA)
```

## Pre-process

```{r}
# pre-process the csv file
species_data <- read.delim("species.csv", sep = "\t", header = TRUE)
species_data <- species_data[, colSums(is.na(species_data)) < nrow(species_data)]
```

```{r}
# Select the key columns for spatial analysis
species_subset <- species_data[, c("scientificName", "decimalLongitude", "decimalLatitude", "coordinateUncertaintyInMeters", "year", "month", "day")]

# Filter out records with missing coordinates
species_clean <- species_subset[!is.na(species_subset$decimalLongitude) & !is.na(species_subset$decimalLatitude), ]
```

## Begin Analysis

```{r}
window <- DATA$Window
class(DATA$Window)
window_sf <- st_as_sf(window)
window_owin <- as.owin(window_sf)
```

## PPP-Object

```{r}
# Convert your species coordinates to match the BC window projection
library(sp)

# Create a SpatialPoints object with your species data
species_sp <- SpatialPoints(
  coords = data.frame(x = species_clean$decimalLongitude, y = species_clean$decimalLatitude),
  proj4string = CRS("+proj=longlat +datum=WGS84")  # GBIF typically uses WGS84
)

# Transform to match the projection of your BC window
species_transformed <- spTransform(species_sp, CRS(proj4string(window)))

# Extract the transformed coordinates
transformed_coords <- coordinates(species_transformed)

# Create the ppp object with transformed coordinates
x_coordinates <- transformed_coords[,1]
y_coordinates <- transformed_coords[,2]

# make a df for the species
species_df <- data.frame(x = x_coordinates, y = y_coordinates)

# Check which points are inside the owin window
inside <- inside.owin(x_coordinates, y_coordinates, window_owin)

# filter the points out that match the window
species_df <- species_df[inside, ]

# create the ppp-object
species_ppp <- ppp(
  x = species_df$x,
  y = species_df$y,
  window = window_owin
)

# plot the window with points
plot(species_ppp, main = "Species Distribution in BC")
```

**Conclusion:**

First impression =\> the distribution is not homo.

## 3D Map with Elevation

```{r}
# get the elevation from DATA
elevation <- DATA$Elevation
figure_3d <- persp(elevation,
      # horizontal and vertical rotation.
      theta = 45, 
      phi = 20,
      # z-axis expansion (need to be bigger for the BC)
      expand = 25,
      # remove border
      border = NA,
      # add a base
      apron = TRUE,
      # shade intensity
      shade = 0.4,
      # axes off.
      box = FALSE,
      main = "elevation of the great blue heron",
      # allow the overlaying for the parks location.
      visible = TRUE,
      # use the terrain color pallete which is built-in for R.
      colmap = terrain.colors(289))

# add the points to the persp plot.
perspPoints(species_df, Z = elevation, M = figure_3d, pch = 16, cex = 0.5)
```

## Elevation class for the specie

```{r}
plot(cut(elevation, 5, labels = c("low", "median-low", "median", "median-high", "high")), main = "elevation classes for the bc blue heron")

# add points
points(species_df$x, species_df$y, pch = 16, col = "black")
```

## Group by distance

```{r}
distances <- nndist(species_ppp)
species_ppp_with_marks <- species_ppp
marks(species_ppp_with_marks) <- data.frame(distance = distances)
col_pal <- colorRampPalette(c("blue", "red"))(100)
dist_scaled <- cut(distances, breaks = 100, labels = FALSE)
point_cols <- col_pal[dist_scaled]
plot(species_ppp, main = "", use.marks = FALSE, cex = 0.2)
title("Group by distance")
points(species_ppp$x, species_ppp$y, pch = 16, col = point_cols)
text(species_ppp$x, species_ppp$y, labels = round(distances/10000, 0), pos = 3, offset = 0.5, cex = 0.7)
```

**Observation:**

Most points are `dark blue` to `purple`, indicating the nearest neighbors are very close. Therefore, the specie tend to cluster in the south of the BC, and most of them are very close. However, the `red` points means that they are far away from their neighbors with `15`, indicating 150 km away from the neighbor.

## Elevation Analysis

```{r}
library(raster)
# 1. Convert spatstat::im (elevation) to raster
elev_raster <- raster(DATA$Elevation)
# get the elevation for specie
elevation_sp <- extract(elev_raster, cbind(species_df$x, species_df$y))
# get the median elevation from the park locations.
median_ele_sp <- median(elevation_sp, na.rm = TRUE)
cat("Median elevation of parks in BC is: ", median_ele_sp)
```

**Conclusion:**

We can see the specie reasonably live in the relatively low elevation area like `low-lying, flat areas`.

```{r}
kde_sp <- density(elevation_sp, na.rm = TRUE)
# plot the elevation density for the specie for better visualization
plot(kde_sp, main = "KDE for blue heron elevations")
```

```{r}
# Extract elevation values from the raster for the whole window (entire raster)
elev_all <- values(elev_raster)

# Extract elevation only at point locations
library(sp)
species_sp <- SpatialPoints(
  coords = data.frame(x = species_ppp$x, y = species_ppp$y),
  proj4string = CRS(projection(elev_raster))
)
elev_points <- extract(elev_raster, species_sp)

# Plot overlaid histograms
hist(elev_all, col = rgb(0, 0, 1, 0.25), main = "Elevation Histogram",
     xlab = "Elevation (m)", border = "white", breaks = 30)
hist(elev_points, col = rgb(1, 0, 0, 0.5), add = TRUE, border = "white", breaks = 30)
legend("topright", legend = c("Whole BC window", "Species locations"),
       fill = c(rgb(0,0,1,0.25), rgb(1,0,0,0.5)), border = NA)
```

**Conclusion:**

Enhance the idea that **most of the specie in the low or median elevation area.**

## Kernel-Density Distribution

```{r}
density_map <- density(species_ppp) 
plot(density_map, main = "Kernel Density of Species Observations")
points(species_ppp, pch = 21, cex = 0.5, col = "black", bg= "white")
contour(density_map, add = TRUE)
```

**Conclusion:**

Majority in the south corner and only few in the north.

## Hot Spot

```{r}
R<-bw.ppl(species_ppp)
LR<-scanLRTS(species_ppp,r=R)
plot(LR)
```

**Observation:**\
Hot spot not distinct at all, maybe because of the number of records is not sufficient. However, still have little colored points in the south indicate that if we want to see the `blue heron`, better go to the south of BC.

## P-value plot

```{r}
p_values <- eval.im(pchisq(LR,df = 1,lower.tail = FALSE))
#Plot the output
plot(p_values, main = "Local p-values For Blue heron")
```

## rho plot

```{r}
#Extract elevation information
elev <- DATA$Elevation
#define quartiles
b <- quantile(elev, probs = (0:4)/4, type = 2)
#Split image into 4 equal-area quadrats based on elevation values
Zcut <- cut(elev, breaks = b)
V <- tess(image = Zcut)
quadratcount(species_ppp, tess = V)
```

**Apparently, most of them (8848) are located in low elevation**

```{r}
rho <- rhohat(species_ppp, elev)
plot(rho)
```

## Quadrat Test

```{r}
# make 10*10 quadrat
Q <- quadratcount(species_ppp, nx = 10, ny = 10)
# quadrat test
quadrat.test(Q)
```

**Conclusion:**

`p-value` small enough, so reject the null. **The assumption about homo is not met.**

```{r}
Q <- quadratcount(species_ppp,
                  nx = 10,
                  ny = 10)

plot(species_ppp,
     pch = 16,
     cex = 0.5,
     cols = "#046C9A",
     main = "Beilschmiedia pendula locations")

plot(Q, cex = 0.5, col = "red", add = T, font = 2)
```

```{r}
K <- Kest(species_ppp, correction = "border")
plot(K, main = "Ripley's K-Function")
```
